The Spine of American Law: Digital History Methods and U.S. Legal Practice

### 2016-07-28 article/Funk-Mullen.Spine-of-Legal-Practice.Rmd
Context of codification

Codification is, as Lawrence Friedman has written, "one of the set pieces of American legal history." Law reformers advocated for a codification of the common law from the earliest days of the Republic through the Gilded Age, from Massachusetts down to South Carolina. Efforts ranged from mere compilations of existing statutes in each state to a full European-style codification meant to be an entirely comprehensive and systematic statement of the law.[^10] At its most basic level, codification proposed that legislative policy ought to be the sole source of law. Law was to be made by democratically responsible legislators in terse, unambiguous statements, not discovered through application and analogy in particular cases by the judges. Debates over codification thus ranged from the metaphysics of law to political theories of institutional competency and the separation of powers.[^11]

@@

[^10]: Friedman, *A History of American Law*, 302. When the intellectual historian Perry Miller developed a reader surveying *The Legal Mind in America*, codification was its only central theme, as Miller argued it was the one intellectual topic that attracted lawyers away from their practices long enough to debate. Perry Miller, ed., *The Legal Mind in America: From Independence to the Civil War* (Anchor 1962). See also Charles M. Cook, *The American Codification Movement: A Study in Antebellum Legal Reform* (1983); Robert W. Gordon, "The American Codification Movement," *Vanderbilt Law Review* 36 (1983): 431--458; Maurice Eugen Lang, *Codification in the British Empire and America* (Lawbook Exchange, 1924). On common theories of codification that transcended jurisdictional boundaries, see Csaba Varga, *Codification as a Socio-Historical Phenomenon*, 2nd ed. (Budapest, 2011 \[1991\]); Roger Berkowitz, *The Gift of Science: Leibniz and the Modern Legal Tradition* (Fordham, 2010).

@@

>>>>

Codification is, as Lawrence Friedman has written, "one of the set pieces of American legal history." Law reformers advocated for a codification of the common law from the earliest days of the Republic through the Gilded Age, from Massachusetts down to South Carolina. Efforts ranged from mere compilations of existing statutes in each state to a full European-style codification meant to be an entirely comprehensive and systematic statement of the law.[^10] At its most basic level, codification proposed that legislative policy ought to be the sole source of law. Law was to be made by democratically responsible legislators in terse, unambiguous statements, not discovered through application and analogy in particular cases by the judges. Debates over codification thus ranged from the metaphysics of law to political theories of institutional competency and the separation of powers.[^11] Codification remained a major interest of the American bar across the nineteenth century. When the intellectual historian Perry Miller developed a reader surveying The Legal Mind in America, codification was its central theme, as Miller argued it was the only intellectual topic that attracted lawyers away from their practices long enough to debate.[^112] The importance of codification extended far beyond the United States. Napoleonic France promulgated a series of codes in the early nineteenth century, Germans debated the wisdom of codification after Napoleon’s fall, and in the 1860s the British imposed codes on their colonies in India and Singapore. Many American lawyers followed the international development of these codes with interest, viewing codification as the leading edge of modern legal science.[^113]

@@

[^10]: Friedman, *A History of American Law*, 302. See also Charles M. Cook, *The American Codification Movement: A Study in Antebellum Legal Reform* (1983); Robert W. Gordon, "The American Codification Movement," *Vanderbilt Law Review* 36 (1983): 431--458; Maurice Eugen Lang, *Codification in the British Empire and America* (Lawbook Exchange, 1924).

@@

[^112]: Perry Miller, ed., *The Legal Mind in America: From Independence to the Civil War* (Anchor 1962).

[^113]: See James Q. Whitman, *The Legacy of Roman Law in the German Romantic Era: Historical Vision and Legal Change* (Princeton, 1990); Gunther A. Weiss, "The Enchantment of Codification in the Common-Law World, *Yale Journal of International Law* 25 (2000): 435; Maurice Eugen Lang, *Codification in the British Empire and America* (Lawbook Exchange, 1924); Jean-Louis Halperin, *The French Civil Code* (Austin: University of Texas Press, 2006); Robert B. Holtman, *The Napoleonic Revolution* (Baton Rouge: Louisiana State University Press, 1981); R. H. Kilbourne, *A History of the Louisiana Civil Code* (Paul M. Herbert Law Center, 1987); Brian Young, *The Politics of Codification: The Lower Canadian Civil Code of 1866* (Osgoode Society, 1994); John W. Cairns, *Codification, Transplants, and History: Law Reform in Louisiana (1808) and Quebec (1866)* (Clark, NJ: Talbot, 2015); Roscoe Pound, "The French Civil Code and the Spirit of Nineteenth Century Law," *Boston Law Review* 35 (1955): 79. On common theories of codification that transcended jurisdictional boundaries, see Csaba Varga, *Codification as a Socio-Historical Phenomenon*, 2nd ed. (Budapest: Akadémiai Kiadó, 2011 [1991]); Roger Berkowitz, *The Gift of Science: Leibniz and the Modern Legal Tradition* (New York: Fordham University Press, 2010).
